---
title: "Kotzebue Phytoplankton Community Composition in September-October 2019"
author: "Mariam Moreno"
date: "13 May 2024"
output:
  pdf_document: default
  html_document: default
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(dplyr)
library(ggplot2)
library(readr)
```


### Aim 2: Enumerate and Identify Phytoplankton

Null hypothesis: Each site will have a different dominant phytoplankton species.

Alternative hypothesis: Both sites will have the same dominant phytoplankton species.

The data is displayed as stacked barplots, one for each site. Different colors represent different phytoplanton identified from lagoon and shore samples from 2019. I anticipate to see each sample site to have a different phytoplankton species dominate the community composition given the differences in locations (the lagoon is closed off and the shore is coastal).
I will use an ANOVA test to determine if both sites have different or the same dominant phytoplankton species.



```{r}
kotz_samples <- read.csv("kotz_samples.csv")

kotz_samples
```


```{r}
# stacked barplot of lagoon and shore counts
kotz_community <- ggplot(kotz_samples, aes(fill=phytoplankton,y=counts, x=site))+ 
  geom_bar(position="stack", stat="identity", colour="black")+ 
  ggtitle("Phytoplankton Counts at Kotzebue Lagoon & Shore")

kotz_community
```



```{r}
# filter data by location
lagoon <- data.frame(filter(kotz_samples, site=="Kotzebue Lagoon"))
lagoon1 <- aggregate(counts~phytoplankton, data=lagoon,sum) 
lagoon1

# finding max of phytoplankton populations
summary(lagoon1)

shore <- data.frame(filter(kotz_samples, site=="Kotzebue Shore"))
shore1 <- aggregate(counts~phytoplankton, data=shore,sum)
shore1

summary(shore1)
```

### ANOVA analysis
```{r}
two.way <- aov(counts~phytoplankton+site, data=kotz_samples)

summary(two.way)
```
I reject my null hypothesis b/c of a very small (insignificant) p-value, and both sites have the same dominant phytoplankton species.
